Manuscript Figures for Genome Biology (simulation subset)

### 2018-08-30 manuscript_figures/manuscript_figures_sims.Rmd

---
title: "Simulation Figures for Genome Biology (simulations)"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

```{r Figure2-load-data}
# Load packages and source benchmark FDR
library(tidyr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(cowplot)
library(tibble)
library(ggthemes)
library(grid)
library(SummarizedBenchmark)

@@

# assumes sb objects are in the following location
path <- file.path("..", "datasets", "simulations", "results")

@@

# Data Preparation

First, we load summaries for all simulations results at significance thresholds
between 0.01 and 0.10. This process can take a long time (approximately 40 minutes
per alpha cutoff) because of the large number of simulation settings that must be
read in and summarized. Ideally, this initial step should be run externally separate
from this Rmd (by default, the following code chunk will not be run).

```{r, eval = FALSE}
for (ia in c(1:10, 20, 30, 40, 50)) {
    summary_file <- paste0("allres_alpha", ia, ".rds")
    if (!file.exists(summary_file)) {
        res_files <- list.files(path, "\\.rds$", full.names = TRUE)
        res <- lapply(res_files, function(x) {
            zz <- readRDS(x)
            zz_i <- lapply(zz, `[[`, "informative")
            zz_u <- lapply(zz, `[[`, "uninformative")
            zz_i <- plotsim_standardize(zz_i, alpha = ia / 100L)
            zz_i <- dplyr::select(zz_i, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
            zz_u <- plotsim_standardize(zz_u, alpha = ia / 100L)
            zz_u <- dplyr::select(zz_u, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
            dplyr::full_join(zz_i, zz_u,
                             by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
                             suffix = c(".info", ".uninfo"))
        })
        names(res) <- gsub("\\.rds$", "", basename(res_files))
        saveRDS(res, summary_file)
    }
}
```

Assuming that the summaries have been saved for all simulation settings at
the various FDR thresholds using the code above, we load the summarized results
and merge them into a single results table.

```{r}
alphas <- 1:10
res <- vector("list", length(alphas))
for (ia in seq_len(length(alphas))) {
    imethodset <- methodset
    imethodset[which(imethodset == "ihw")] <- paste0("ihw-a", sprintf("%02d", alphas[ia]))
    ires <- readRDS(paste0("allres_alpha", alphas[ia], ".rds"))
    ires <- bind_rows(ires, .id = "setting")
    ires <- dplyr::mutate(ires, dist = gsub(".*?-benchmark-(.*)", "\\1", setting),
                         setting = gsub("(.*?)-benchmark-.*", "\\1", setting))
    ires <- dplyr::filter(ires, blabel %in% imethodset)
    ires <- dplyr::mutate(ires, blabel = gsub("(-a)(.*)", "", blabel))
    res[[ia]] <- ires
}
res <- bind_rows(res)
res <- dplyr::select(res, -param.alpha)
res <- dplyr::rename(res, Method = blabel)
res <- dplyr::mutate(res, Method = gsub("-df03", "", Method))
res <- dplyr::mutate(res, Method = factor(Method, levels = gsub("-df03", "", methodset)))
```

We compute summary metrics for each method, setting, and cutoff.

```{r}
resmet <- dplyr::group_by(res, setting, dist, Method, key, performanceMetric, alpha)
resmet <- dplyr::summarize(resmet,
                           mean.info = mean(value.info, na.rm = TRUE),
                           se.info = sd(value.info, na.rm = TRUE) / sqrt(sum(!is.na(value.info))),
                           nNA.info = sum(is.na(value.info)),
                           mean.diff = mean(value.info - value.uninfo, na.rm = TRUE),
                           se.diff = sd(value.info - value.uninfo, na.rm = TRUE) /
                               sqrt(sum(!is.na(value.info - value.uninfo))))
resmet <- dplyr::ungroup(resmet)
```

We clean up the annotations in the table for some of the more complex simulation settings.

```{r}
resmet <- dplyr::mutate(resmet, inform = ifelse(grepl("varyinginfo", setting), dist, NA))
resmet <- dplyr::mutate(resmet, inform = as.numeric(gsub("level", "", inform)))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyinginfo", setting), "gaussian", dist))

resmet <- dplyr::mutate(resmet, pi0 = ifelse(grepl("varyingpi0", setting), dist, 90))
resmet <- dplyr::mutate(resmet, pi0 = as.numeric(gsub("nullprop", "", pi0)))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyingpi0", setting), "gaussian", dist))

resmet <- dplyr::mutate(resmet, signal = ifelse(grepl("varyinges", setting), dist, NA))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges$", setting), "gaussian", dist))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-t$", setting), "t11", dist))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-chisq$", setting), "chisq4", dist))
resmet <- dplyr::mutate(resmet, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
```

We also replace the sampling noise labels with more plotting-friendly labels.

```{r}
resmet <- dplyr::mutate(resmet, dist = factor(dist, levels = c("gaussian", "t11", "t5", "chisq4"),
                                              labels = c(expression("Noise: " * N(0, 1)),
                                                         expression("Noise: " * t[11]),
                                                         expression("Noise: " * t[5]),
                                                         expression("Noise: " * {chi^2}[4]))))
```

@@

There are several different simulation settings that can be investigated.

```{r}
dplyr::distinct(resmet, setting)
```

Namely, simulation settings were run with to investigate the impact of different
informative covariates, noninformative covariates, varying null proportions,
varying covariate informativeness, and varying effect size distributions (including
unimodal effects as described in the ASH manuscript). Many of these simulation
settings were also run with different sampling noise distributions, including
Normal/Gaussian, t with 11 degrees of freedom, t with 5 degrees of freedom, and
Chi-squared with 4 degrees of freedom.

# Null Simulations

We subset on simulation results that look at informative and non-informative covariates across
different noise distributions.

For null simulations, rather than the false discovery rate, we take a look at the true negative
rate (equivalently, the proportion of false positives) since the FDR will be 1 if any rejections
are made. Additionally, we take a look at the FWER (the proportion of simulation replicates out
of 100 with at least one rejection). In these plots, `ashq`, `fdrreg-e` and `fdrreg-t`
are excluded for the Chi-Squared simulation settings.

```{r, fig.width = 18, fig.height = 8}
res_nullset <- dplyr::filter(resmet, setting == "null",
                             !(Method == "ashq" & grepl("chi\\^2", dist)),
                             !(grepl("fdrreg", Method) & grepl("chi\\^2", dist)))
res_nullset <- tidyr::complete(res_nullset, setting, Method, dist, key)

null_tnr <- dplyr::filter(res_nullset, key == "TNR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TNR", labels = scales::percent) +
    ggtitle("Mean TNR Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

null_fwer <- dplyr::filter(res_nullset, key == "FWER") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FWER", labels = scales::percent) +
    ggtitle("FWER Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

gp <- plot_grid(null_tnr + guides(color = FALSE, linetype = FALSE),
                null_fwer + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(null_tnr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
null_tnr_diff <- dplyr::filter(res_nullset, key == "TNR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TNR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TNR Difference Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

null_fwer_diff <- dplyr::filter(res_nullset, key == "FWER") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FWER\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("FWER Difference Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

gp <- plot_grid(null_tnr_diff + guides(color = FALSE, linetype = FALSE),
                null_fwer_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(null_tnr_diff), rel_widths = c(1, 1/16))
gp
```

# Sampling Distributions

We just take a look at different noise distributions for a single informative covariate.

```{r, fig.width = 18, fig.height = 8}
res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
res_noise <- tidyr::complete(res_noise, setting, Method, dist, key)

noise_fdr <- dplyr::filter(res_noise, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

noise_tpr <- dplyr::filter(res_noise, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
noise_fdr_diff <- dplyr::filter(res_noise, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

noise_tpr_diff <- dplyr::filter(res_noise, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

gp <- plot_grid(noise_fdr_diff + guides(color = FALSE, linetype = FALSE),
                noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
gp
```

# Informative Covariates

We also take a look across different informative covariates.

```{r, fig.width = 18, fig.height = 8}
res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
res_icov <- tidyr::complete(res_icov, setting, Method, dist, key)

icov_fdr <- dplyr::filter(res_icov, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

icov_tpr <- dplyr::filter(res_icov, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

gp <- plot_grid(icov_fdr + guides(color = FALSE, linetype = FALSE),
                icov_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(icov_tpr), rel_widths = c(1, 1/12))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
icov_fdr_diff <- dplyr::filter(res_icov, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

icov_tpr_diff <- dplyr::filter(res_icov, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

gp <- plot_grid(icov_fdr_diff + guides(color = FALSE, linetype = FALSE),
                icov_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(icov_tpr_diff), rel_widths = c(1, 1/12))
gp
```

# Noninformative Covariates

We take a look at noninformative covariates on their own.

```{r, fig.width = 18, fig.height = 8}
res_noni <- dplyr::filter(res, grepl("informative-cubic", setting))
res_noni <- dplyr::group_by(res_noni, setting, dist, Method, key, performanceMetric, alpha)
res_noni <- dplyr::summarize(res_noni,
                             mean.uninfo = mean(value.uninfo, na.rm = TRUE),
                             se.uninfo = sd(value.uninfo, na.rm = TRUE) /
                                 sqrt(sum(!is.na(value.uninfo))))
res_noni <- dplyr::ungroup(res_noni)

noni_fdr <- dplyr::filter(res_noni, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.uninfo, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.uninfo - se.uninfo, ymax = mean.uninfo + se.uninfo), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications (noninformative covariates)") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

noni_tpr <- dplyr::filter(res_noni, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.uninfo, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.uninfo - se.uninfo, ymax = mean.uninfo + se.uninfo), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications (noninformative covariates)") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

gp <- plot_grid(noni_fdr + guides(color = FALSE, linetype = FALSE),
                noni_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noni_tpr), rel_widths = c(1, 1/16))
gp

gp <- plot_grid(noni_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
                noni_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noni_tpr), rel_widths = c(1, 1/16))
gp
```

# Null Proportions

We take a look at the impact of the proportion of null hypotheses.

```{r, fig.width = 10, fig.height = 4}
res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0-lowsignal",
                         alpha == 0.05)
res_pi0 <- dplyr::mutate(res_pi0,
                         mean.info = ifelse(is.nan(mean.info) & key == "FDR", 0, mean.info))

pi0_fdr <- dplyr::filter(res_pi0, key == "FDR") %>%
    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

pi0_tpr <- dplyr::filter(res_pi0, key == "TPR") %>%
    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE),
                pi0_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 2)
gp <- plot_grid(gp, get_legend(pi0_tpr), rel_widths = c(1, 1/8))
gp
```

# Informativeness

We take a look at the impact of the informativeness of a covariate.

```{r, fig.width = 10, fig.height = 4}
res_info <- dplyr::filter(resmet, setting == "varyinginfo-lowsignal", alpha == 0.05)

info_fdr <- dplyr::filter(res_info, key == "FDR") %>%
    ggplot(aes(x = inform, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

info_tpr <- dplyr::filter(res_info, key == "TPR") %>%
    ggplot(aes(x = inform, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

gp <- plot_grid(info_fdr + guides(color = FALSE, linetype = FALSE),
                info_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 2)
gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, 1/8))
gp
```

We also look at the difference between informative and uninformative covariates.

>>>>

---
title: "Manuscript Figures for Genome Biology (simulation subset)"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

Running this

```{r Figure2-load-data}
## Load packages and source benchmark FDR
library(tidyverse)
library(cowplot)
library(ggthemes)
library(grid)
library(SummarizedBenchmark)

@@

# assumes sb objects are in the following location
resmet_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")

@@

@@

To generate the figures in this document, the simulation results must first be aggregated by running the code at `datasets/simulations/simulations-summary.Rmd`.

```{r}
resmet <- readRDS(resmet_file)
```

Unfortunately, because the plots in this section illustrate FDR and TPR acorss varying simulation settings rather than varying nominal FDR cutoff, we cannot simply use the `plotsim_average()` function to generate plots. Instead, we define the following helper function to consistently generate plots similar to those output by `plotsim_average()` with arbitrary covariate, e.g. number of tests or proportion of null hypotheses, as the x-axis. The function assumes that the input table only includes FDR or TPR values for each method at a single nominal alpha cutoff. In these figures, we only plot FDR and TPR values at the nominal alpha cutoff of 0.05.

```{r}
genplot <- function(tab, cov, xt = "", met = c("FDR", "TPR"), ebw = 0.0025) {
    met <- match.arg(met)
    cov <- rlang::enquo(cov)
    gp <- dplyr::filter(tab, key == met) %>%
        ggplot(aes(x = !!cov, y = mean.info, color = Method, group = Method)) +
        geom_line(aes(linetype = Method), alpha = 0.85) +
        geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = ebw, alpha=0.5) +
        scale_linetype_manual(values = lty) + 
        scale_color_manual(values = col) +
        scale_y_continuous(met, labels = scales::percent) +
        expand_limits(y = 0) +
        ggtitle(paste("Mean", met)) +
        theme_classic() +
        theme(axis.title = element_text(face = "bold", size = 10),
              plot.title = element_text(face = "bold", size = 14))
    if (met == "FDR") {
        if (rlang::quo_name(cov) == "alpha") {
            gp <- gp + geom_abline(lty = 2, color = "blue", alpha = 1/2) 
        } else {
            gp <- gp + geom_hline(yintercept = 0.05, lty = 2, color = "blue", alpha = 1/2) 
        }
    }
    gp
}
```

# Figure 3

This is a faceted figure with results from the pure simulation studies showing several key results from a subset of simulation settings. Namely, the figure includes plots of FDR and TPR values at a nominal FDR cutoff of 0.05 over 100 replications illustrating the impact of (A,B) changing total number of tests, (C,D) changing the proportion of non-null tests, and (E,F) changing the informativeness of a covariate.

```{r, fig.width = 10, fig.height = 12}
## Figure 3A,B: FDR/TPR of methods across varying number of tests at alpha = 0.05
res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
p3a <- genplot(res_ntests, cov = as.factor(ntests), met = "FDR", ebw = 0.15) +
    xlab("Number of Tests")
p3b <- genplot(res_ntests, cov = as.factor(ntests), met = "TPR", ebw = 0.15) +
    xlab("Number of Tests")

## Figure 3C,D: FDR/TPR of methods across varying null proportion at alpha = 0.05
res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
p3c <- genplot(res_pi0, cov = pi0, met = "FDR", ebw = 2.5) +
    scale_x_continuous("Proportion of Null Hypotheses",
                       breaks = c(5, seq(10, 90, by = 10), 95, 99))
p3d <- genplot(res_pi0, cov = pi0, met = "TPR", ebw = 2.5) +
    scale_x_continuous("Proportion of Null Hypotheses",
                       breaks = c(5, seq(10, 90, by = 10), 95, 99))

## Figure 3E,F: FDR/TPR of methods across varying informativeness at alpha = 0.05
res_info <- dplyr::filter(resmet, setting == "varyinginfo", alpha == 0.05)
p3e <- genplot(res_info, cov = inform, met = "FDR", ebw = 2.5) +
    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))
p3f <- genplot(res_info, cov = inform, met = "TPR", ebw = 2.5) +
    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))

Fig3 <- plot_grid(p3a + guides(color = FALSE, linetype = FALSE),
                p3b + guides(color = FALSE, linetype = FALSE),
                p3c + guides(color = FALSE, linetype = FALSE),
                p3d + guides(color = FALSE, linetype = FALSE),
                p3e + guides(color = FALSE, linetype = FALSE),
                p3f + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:6], ncol = 2)
Fig3 <- plot_grid(Fig3, get_legend(p3a), rel_widths = c(1, .2))
Fig3
ggsave(file.path(outdir, "Figure3.pdf"), width=10, height=12)
```

# Figure 3+

This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different test statistic distributions. Test statistics were simulated from Normal, t with 11 degrees of freedom, t with 5 degrees of freedom, and Chi-squared with 4 degrees of freedom.

```{r, fig.width = 10, fig.height = 12}
## Figure 3plus A,B: FDR/TPR of methods across varying test statistics distribution
res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
p3pa <- genplot(res_noise, cov = alpha, met = "FDR") +
    facet_grid(. ~ dist, labeller = label_parsed)
p3pb <- genplot(res_noise, cov = alpha, met = "TPR") +
    facet_grid(. ~ dist, labeller = label_parsed)

Fig3p <- plot_grid(p3pa + guides(color = FALSE, linetype = FALSE),
                   p3pb + guides(color = FALSE, linetype = FALSE),
                   labels = LETTERS[1:2], ncol = 1)
Fig3p <- plot_grid(Fig3p, get_legend(p3pa), rel_widths = c(7.5, 1))
Fig3p
ggsave(file.path(outdir, "Figure3plus.pdf"), width=14, height=8)
```

# Session information

```{r}
sessionInfo()
```
